Learning and comparing functional connectomes across subjects

### 2013-01-01 paper.tex
Iterate on the predictive modeling part

\cite{craddock2009}
craddock 2009, Dosenbach 2010, lord 2011, deshpande,

1. importance of interpretation!
2. the need for feature extraction
3. the need for feature selection
4. setting parameters.
5. different methods have different interpretations

@@

% Feature extraction and parameter selection

@@

Predictive modeling has been an active area of research in functional
neuroimaging since the early 90s [] that has more recently been adapted to
comparing connectomes [].  A predictive model defines a mathematical
relationship between a set of data features and a categorical or continuous
variable of interest (target) that is capable of decoding or \emph{predicting}
the target variable from the data.  In the context of comparing connectomes,
features correspond to bivariate measures of connectivity or other region
specifc graph statistics and the target is a phenotypic variable (e.g. age, sex,
disease state). The accuracy of a model for predicting the target variable
provides a natural means for evaluating the utility of the model as a biomarker
[] and for measuring the amount of information present in the connectome about
the target variable [Kjems, Kriegeskorte]. Prediction accuracy when combined
with a measure of model reproducibilty provides a means to evaluate experimental
tradeoffs for data acquisition, preprocessing, and analysis [NPAIRS]. Predictive
modeling techniques have been successfully applied to identify connectome-based
biomarkers of Alzheimer's disease [92], major depression [93], schizophrenia
[94], autism [47], ADHD [], and aging [].  Recent work has illustrated the
utility of predictive modeling for deriving connectivity models at the
individual level [97,].

Nearly every analytical method can be applied in a predictive modeling framework, but
multivariate classification and regression have been the most commonly used
(referred to as MVPA; multivariate prediction analysis).  Multivariate methods
are attractive in connectomics because they are sensitive to dependencies
between features and avoid the need to correct for multiple comparisions since
the significance of an entire pattern is evaluated using a single statistical
test. Predictive modeling is typically performed in a supervised learning
framework where a target is available for each observation of the features,
although unsupervised and semi-supervised methods can also be used.

Predictive models are validated using one or more iterations of cross
validation, which involves training and testing phases.  A cross validation
cycle begins by seperating the available data into two independent subsets:
training data and testing data.  Training involves learning a predictive model
from the training data. During testing the model is applied to the testing data
to predict corresponding values of the target variable.  Prediction accuracy is
then assessed by comparing these predictions to the observed values of the
target variable. In order to make

A general application of predictive modeling can be broken down into feature
extraction, model selection, feature selection, and model validation, each of
which may contain one or more iteration of training and testing. Feature
extraction

Model training can be further broken down into feature extraction, model
selection and feature selection and testing can be performed for model
optimization and

The application of predictive modeling to functional neuroimaging data has been
an active area of research since the early 90s[] and more recently for comparing
connectomes[]. Nearly every analytical method can be appplied in a predictive
modelling framework, but multivariate classification and regression have been
the most commonly used (referred to as MVPA; multivariate prediction analysis).
Multivariate methods are attractive in connectomics because they do not require
any assumptions regarding independence between features and avoid the need to
correct for multiple comparisions since the significance of an entire pattern is
evaluated using a single statistical test. Support vector machines
(classification and regression) are popular for MVPA applications owing to their
proven track record for acheiving good prediction accuracy in high dimensional
datasets (many features) with few observations.

Cross validation

MVPA algorithms typically require the specification of several parameters, the
values of which will impact the quality of the results. These parameters may be
prespecified based on domain specific knowledge, or requirements, determined
using an analytical approach, or optimized using cross validation. In latter
case a nest cross-validation scheme is required to avoid biasing resultant
measures of prediction accuracy. it is necessary to perform a nested
cross-validation These parameters may be set by hand in the case where

Determining the subset of features most relevant to the model (feature
selection) is essential for interpreting the models, and is the subject of
ongoing research. In cases were model weights can be directly mapped to their
corresponding feature, there is typically no statistical framework for
thresholding these weights, and meaningless features are not always set to zero
(this is particularly the case for L2 regularized algorithms such as SVM and
ridge regression). One feature selection strategy is to filter out features
based on their statistical relationship with the variable of interest using a
univariate statistical criterion such as t-test, Pearson's correlation, or
mutual information.  A drawback of this approach is that since the filters are
typically insensitive to multivaratiate information, they may exclude features
that would be important to the model.  Another approach is to treat feature
selection as a model selection problem in which case the features are
iteratively pruned and a subset is chosen based on prediction performance. A
third approach is to use sparse modeling methods such as L1 SVM or LASSO that
force meaningless weights to be zero.  Although ideal, these methods may have
limits on the number of non-zero weights used in the model or the exclusion of
highly correlated features from the model.  Contrary to previous reports [Guyon
2002, Craddock 2009] no matter which feature selection strategy is employed, it
must be performed with cross-validation to avoid biasing estimates of
prediction accuracy [later Guyon].

The interpretation of classifier weights for selected features, when available,
is complicated and requires insight into the mechanisms underlying the modeling
approach. If features are appropriately standardized prior to training, the
magnitude of weights can be interpreted to reflect the relative importance of
the feature to the model, but interpreting how the feature differs between
classes or relates to a continuous variable is more complicated, particularly
given the multivariate nature of their involvement.  For classification methods
such as Fisher's linear discriminant analysis (LDA), which maximally seperates
class means given their standard deviation, feature weights are representative
of the differences between class means. For support vector classification,
classifier weights are derived from observations that are the most similar
between classes, and hence represents the border between classes.  It is perhaps
most reasonable to adapt a conservative interpretation in which MVPA is used to
identify the most relevant features that are later tested in followon
experiments better suited to elucidating their relationship to the variable of
interest.

@@

It is tempting to favor more neurobiologically-inspired models that give
descriptions close to our knowledge of the brain basic mechanisms,
however, as George Box famously said, ``all models are wrong; some models
are useful''. Depending on the question and the data at hand, the cursor
should be set between complex models based on a bio-physical description,
and simple phenomenological models such as correlation matrices. In
particular to model interactions between a large number of regions, as in
full-brain analysis, and learn a large \emph{connectome}, simple models
are to be preferred. For more hypothesis-driven studies, such as the
analysis of the mechanism underlying a specific task, more complex models
can be preferred, if rich data is available. Automatic choice of model is
a difficult problem, however, cross-validation (as used in
\cite{varoquaux2010c,craddock2011,strother2006}) is a useful tool. The
central principle of cross-validation is to learn (or fit) a model, and
then test its predictive power on unseen data. Models too complex
for the data available will fit noise in the data, and thus generalize
poorly. The main benefit of cross-validation is that it is a
non-parametric method which does not rely strongly on modeling
assumptions\footnote{This is to be contrasted to Bayesian model
comparison, which will give correct results only if the true generative
model is in the list of models compared. \cite{friston2012} argues that,
based on the Neyman-Pearson lemma, cross-validation is less powerful than
likelihood ratio tests relying on the full dataset. However, it is
important to keep in mind that these approaches only test for
self-consistence, as the Neyman-Pearson lemma requires that 
the model likelihood describes the data well which may not always be the
case\cite{lohmann2012}.}.

>>>>

@@

Predictive modeling is concerned with learning (or fitting) a model to
draw conclusion on new and unseen data \cite{pereira2009}. Typically on
brain imaging data, it opens the door to learning diagnosis or prognosis
markers \cite{stonnington2010}. However, beyond the practical use of
prediction, these models can also be useful to uncover mechanism of brain
function \cite{norman2006}. Both the fact that information \emph{can} be
drawn to generalize predictions to new data and \emph{how} or
\emph{where} informs us on brain function. Multivariate predictive models
are attractive in connectomics because they are sensitive to dependencies
between features and avoid the need to correct for multiple comparisons
since the significance of an entire pattern is evaluated using a single
statistical test. From the perspective of medical applications, using
resting-state connectivity as a biomarker is particularly appealing as it
can be applied to impaired patients \cite{greicius2008b}. Predictive
modeling techniques have been successfully applied to identify
connectome-based biomarkers of Alzheimer's disease
\cite{stonnington2010}, depression \cite{craddock2009}, schizophrenia
\cite{cecchi2009}, autism \cite{anderson2011}, ADHD \cite{zhu2008}, aging
\cite{dosenbach2010}, as well as to classify mental operations
\cite{richiardi2011,shirer2012}.

% A bit of technical details

Technically, predictive modeling is a supervised machine learning problem
where a target to be predicted --\emph{e.g.}\ age, disease state,
cognitive state-- is available for each observation of the data. In the
context of comparing connectomes, features used in the predictive model
usually correspond to bivariate measures of connectivity, or any other
summary metrics discussed earlier.
%
The prediction accuracy of a model for predicting the target variable
provides a natural means for evaluating the utility of the model as a
biomarker, but also its explanatory power for the mechanism under study.
%
This prediction accuracy is measured using one or more iterations of
cross validation, which involves training and testing phases
\cite{pereira2009}.  A cross validation cycle begins by separating the
available data into two independent subsets: training data and testing
data.  Training involves learning a predictive model from the training
data. During testing the model is applied to the testing data to predict
corresponding values of the target variable.  Prediction accuracy is then
assessed by comparing these predictions to the observed values of the
target variable.

% XXX: sentence below should probably be moved elsewhere
Support vector machines (classification and regression) are popular
algorithms for predictive modeling as they do not require fine parameter
setting.
% XXX: need better explanation, in particular of nested cross-val
However, learning algorithms require in general the specification of
parameters, the values of which will impact the quality of the results.
These parameters may be specified based on domain specific knowledge, or
requirements, determined using an analytical approach, or optimized using
cross validation. In latter case a nest cross-validation scheme is
required to avoid biasing resultant measures of prediction accuracy.

Determining the subset of features most relevant to the model (feature
selection) is essential for interpreting the models. In cases were model
weights can be directly mapped to their corresponding feature, there is
typically no statistical framework for thresholding these weights, and
meaningless features are not always set to zero (this is particularly the
case for $\ell_2$ regularized algorithms such as SVM and ridge
regression). One feature selection strategy is to filter out features
based on their statistical relationship with the variable of interest
using a univariate statistical criterion such as t-test, Pearson's
correlation, or mutual information.  A drawback of this approach is that
since the filters are typically insensitive to multivariate
information, they may exclude features that would be important to the
model.

% XXX: mention nested cross-val for feature selection above

@@

% XXX: distinguish linear from non-linear classifiers. Explain that a
% linear classifier is nothing more than a fancy linear regression

The interpretation of classifier weights for selected features, when
available, is complicated and requires insight into the mechanisms
underlying the modeling approach. If features are appropriately
standardized prior to training, the magnitude of weights can be
interpreted to reflect the relative importance of the feature to the
model, but interpreting how the feature differs between classes or
relates to a continuous variable is more complicated, particularly given
the multivariate nature of their involvement.  For classification methods
such as Fisher's linear discriminant analysis (LDA), which maximally
separates class means given their standard deviation, feature weights are
representative of the differences between class means. For support vector
classification, classifier weights are derived from observations that are
the most similar between classes, and hence represents the border between
classes. To draw conclusions on the features used for prediction, it is
perhaps most reasonable to adapt a conservative interpretation in which
predictive models is used to identify the most relevant features that are
later tested in followon experiments better suited to elucidating their
relationship to the variable of interest.

@@

It is tempting to favor more neurobiologically-inspired models that give
descriptions close to our knowledge of the brain basic mechanisms,
however, as George Box famously said, ``all models are wrong; some models
are useful''. Depending on the question and the data at hand, the cursor
should be set between complex models based on a bio-physical description,
and simple phenomenological models such as correlation matrices. In
particular to model interactions between a large number of regions, as in
full-brain analysis, and learn a large \emph{connectome}, simple models
are to be preferred. For more hypothesis-driven studies, such as the
analysis of the mechanism underlying a specific task, more complex models
can be preferred, if rich data is available. Automatic choice of model is
a difficult problem, however, cross-validation (as used in
\cite{varoquaux2010c,craddock2011,strother2006}) is a useful tool. The
central principle of cross-validation is to learn (or fit) a model, and
then test its predictive power on unseen data. Models too complex
for the data available will fit noise in the data, and thus generalize
poorly. The main benefit of cross-validation is that it is a
non-parametric method which does not rely strongly on modeling
assumptions\footnote{This is to be contrasted to Bayesian model
comparison, which will give correct results only if the true generative
model is in the list of models compared. \cite{friston2012} argues that,
based on the Neyman-Pearson lemma, cross-validation is less powerful than
likelihood ratio tests relying on the full dataset. However, it is
important to keep in mind that these approaches only test for
self-consistence, as the Neyman-Pearson lemma requires that 
the model likelihood describes the data well which may not always be the
case \cite{lohmann2012}.}.
